---
title: 'Cleaning Evaluation Data: Grouping - Summarise - Export'
author: "Corey Pembleton"
date: '2018-11-22'
output: html_document
draft: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Essential R Toolkit in Data Cleaning: Grouping, Summarising and Exporting

I'm yet to participate in an evaluation that doesn't have the standard data science workflow of importing-tidying-understanding-presenting. The tidying and understanding components of this are tools every evaluator should have readily at their finger tips. Being able to reuse the same tools and methods across projects is a smarter way of working, especially when last-minute tables, graphs, presentation material and additional analyses are needed. 

```{r, message=FALSE, warning=FALSE}
###The libraries used for this post
library(tidyverse)
library(openxlsx)
library(kableExtra)
```




```{r, message=FALSE}
```

At a glance, the table is 

```{r, echo=FALSE}

```





#### Reference
Max Roser (2018) - "Child Mortality". Published online at OurWorldInData.org. Retrieved from: 'https://ourworldindata.org/child-mortality' [Online Resource]


