---
title: "Creating a Proportional Random Sample of a Grouped Variable"
author: "Corey Pembleton"
date: '2018-10-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(kableExtra)

```

```{r, echo=FALSE}
Portfolio <- read_excel("Sample_example.xlsx", sheet = 1)
```

In this post, I create a randomized sample based upon the proportional representation of a given group, something which often arises in development monitoring and evaluation studies. An example of this would be when you want to select a certain percentage of projects on a specified group (e.g. thematic area, region, leader responsible) relevant to its proportional size. 

I wasn't so sure how to go about doing this smoothly in a tidy way, until I found [this post](https://jennybc.github.io/purrr-tutorial/ls12_different-sized-samples.html) by Jenny Bryan, which helped guide me along the way, and taught me a few new tricks such as the value of nesting and unnesting dataframes, and using the ```map2()``` purrr function within a tidy workflow.

In the post which this draws from the sample size is determined according to counts of select variable groups. Here I opt for a proportional approach representative of the groups of the population, in this case by regional representation. 

Below, I determine the proportion of the population (country_total) based on a 30% sample size (sample_count).

```{r, results='hide'}
Portfolio %>% 
  group_by(Region) %>% 
  count() %>% 
  summarise(country_total = n,
            prop = n/91,
            sample_count = round(prop*30)) %>%
  arrange(Region)
```

Which doesn't need to be saved, I can just copy the values according to their proportion (in this case, I'm aiming for a sample which is 30% of the total population, by Regional representation):  


```{r, echo=FALSE}
Portfolio %>% 
    group_by(Region) %>% 
  count() %>% 
  summarise(country_total = n,
            prop = n/91,
            sample_count = round(prop*30)) %>%
  arrange(Region) %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(width = "100%", height = "300px")

```

Using the results from the country_sample column, I can then select the random sample itself. This works by extracting a random sample of the desired amount according to the "sample count" value above (so 7 will be selected from Asia, 9 from LAC, 13 from SSA) using a ```purrr::map()``` of a sample using the ```sample_n()``` function. each time this is run, it extracts a new random sample, but always with the same given count per group.

```{r, echo = FALSE}

Portfolio %>% 
  arrange(Region) %>% 
  group_by(Region) %>% 
  nest() %>% 
  mutate(n = c(7,0,9,1,13),
         samp = map2(data, n, sample_n)) %>% 
  select(Region, samp) %>% 
  unnest() %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(width = "100%", height = "300px")

```

And with that, I know have a sample according to each of the 5 region's representative proportion, grouped according to each region's proportion. Each time I run this script, I'll get a new random sample from each region of the count specified. 

The beauty of this approach is its flexibility, and that it is easy to interpret and manipulate according to needs and data sizes. The nesting of groups of rows based upon certain variables has clear advantages, and is something I anticipate to be more inclusive of in my development-evaluation sampling methods going forward. 



